---
title: "COVID 19 Tracking"
author: "Katelyn Navarette"
output: git_hub
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
library(lubridate)
library(geofacet) #install for NY Times Map
library(zoo) #install to deal with time series
library(scales) #can make nicer axis labels for graphs
```

## Setup - United States COVID-19 Cases and Deaths

Here are the NY Times data on cumulative cases by date and state.
```{r}
cases <- read_csv("https://github.com/nytimes/covid-19-data/raw/master/us-states.csv")
```

Note that both `cases` and `deaths` are cumulative by date.

If we want to extract the daily cases we can use the following code which will calculate the difference in (cumulative) cases for one day minus the previous day using the `diff()` function 
```{r}
cases %>%
  group_by(state) %>%
  mutate(case_daily = c(cases[1], diff(cases)),
         death_daily = c(deaths[1], diff(deaths))) %>%
  ungroup()
```


It is also common to report the 7-day moving average to account for reporting delays over the weekends and delays testing results reports.

```{r}
cases %>%
  group_by(state) %>%
  mutate(case_daily = c(cases[1], diff(cases))) %>%
  #filter(!(date == as.Date("2021-03-08") & state == "Missouri")) %>% 
  arrange(date) %>%
  mutate(case_7 = rollmean(case_daily, k = 7, fill = NA)) %>%
  ungroup()
```


Here is data from the US Census Bureau on the state estimated population size. Why might this data be useful?
```{r}
pop <- read_csv("USpop2019.csv")
```


## Part 1 - Identify Research Questions
Identify 2-3 potential research questions you could try to answer using the provided data (or if you had additional data).

>Does population size corrilate to the states area in square miles?  
 
>Whichh state is bwcomeing the most populat to live in (you woudl need the population of the states of ther the past few years and see where theres a increase of decrease in the precent they usually rise up by.)?  

> Is the distribution of government reresentatives even throuhg out the states, or wven which states population is best represented in their state? (aka which state has the most people in poitical involvment)


## Part 2 - Outline the Process
Pick one of the research questions above. Write a brief outline of the steps you would need to take to 1) prepare the data for analysis, and 2) analyze/visualize the data. 

Data Preparation  

- Step 1: I would have to collect the square miles from each state and make sure there was something to bind the data sets together, aka the GEOID. 
- Step 2: Left join the data 
- Step 3: I would also do a count in the case there are any missing states this would check that 
- Step 4: create a new variable showing the population divided by the state square miles to see the proportion for each state



Data Analysis/Visualization  

- Step 1:  create a ggplot showing the x asis being the state and the y being the proportion
- Step 2:  the plat you will make is a histogram which can



## Part 3 - Start Coding
Start to translate your outline into code. You may need to look back at cheat sheets, previous notes, or Google items.  Start simple first and then add complication.

>(this is what the data woudl look like if there was data on the state size)

```{r}
pop %>% 
  left_join(state_size) %>% 
  count(NAME) %>% 
  mutate(prop = (estimate/area)) -> State_Props
  
```



```{r}
State_Props %>% 
  ggplot(aes(x = state, y = prop, fill = state)) +
  geom_histogram()
```


Now take what you have created and think about ways to modify your visualization to make it more interesting or engaging, to clean it up, add details, or highlight details.  

```{r}
 State_Props %>% 
  ggplot(aes(x = state, y = prop, fill = state)) +
  geom_histogram() +
  labs(x = "States",
       y = "Proportion of State Population by State Area",
       title = "State Area cover compaired to State populaiton") +
  theme_minimal() 
```


## Part 4 - Look Back, Look Ahead
Look back at what you have done, were you able to adequately answer your original research question? Look ahead at what steps you might take next to improve on your analysis or to modify your existing visualization (even if you are not sure how to do it in R).


> i noticed i may have to do somethign more to my dadta to get an idea of how the larger states comparer to that of the smaller states im not sure how i woudl graph that but it woudl eb voom ifi ihave the proper data set witht eh state sizes in sqmi. 




## Extra Extra
Recently, the NY Times provided a graph of cases rising across the US (scroll down to "After month of progress...")
https://www.nytimes.com/interactive/2021/04/06/us/variants-cases-spread.html


A fellow statistician recreated the graphic using R:
https://livefreeordichotomize.com/2021/04/07/nytimes-map-how-to/

I have already provided the US Census data so you can skip those steps, but give it a try, replicating her code to see if you can recreate the graphic. Comment the code as you determine what each line does.
